---
title: "Week 6 - Neural Networks"
author: "Dolores Farley - George Mason University, School of Business"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: sentence
---

```{r}
library(neuralnet)
library(nnet)
library(caret)
library(e1071)
library(rsample)
library(fastDummies)
library(tidyverse)
library(factoextra)
library(skimr)
library(dplyr)


EmployeeData <- read_csv("~/Desktop/GMU School/MBA_739/Week6/Assignment6/EmployeeData.csv")
EmployeeData <- EmployeeData %>%
           mutate(left_company = as.factor(left_company))%>%
  mutate(job_satisfaction = as.factor(job_satisfaction))%>%
  mutate(performance_rating = as.factor(performance_rating))%>%
  mutate(department = as.factor(department))%>%
  mutate(job_level = as.factor(job_level))%>%
  mutate(business_travel = as.factor(business_travel))%>%
  mutate(marital_status = as.factor(marital_status))

# Separate the variables to be scaled and the dummy variables
scale_vars <- EmployeeData %>%
  select(-c("job_satisfaction", "performance_rating", "department","job_level", "business_travel", "marital_status","left_company")) 

# Separate the variables to be scaled and the dummy variables

dummy_vars <- EmployeeData %>%
  select(c("job_satisfaction", "performance_rating", "department","job_level", "business_travel", "marital_status","left_company")) 

# Add or remove variables as needed

# Scale the variables
maxs <- apply(scale_vars, 2, max)
mins <- apply(scale_vars, 2, min)


scale_vars <- scale_vars %>%
  scale(center = mins, scale = maxs - mins) %>%
  as.data.frame()


dummy_vars <- dummy_cols(dummy_vars, select_columns=c("job_satisfaction", "performance_rating", "department","job_level", "business_travel", "marital_status"),
           remove_first_dummy = FALSE,
           remove_selected_columns = TRUE)

employeeDF <- bind_cols(scale_vars, dummy_vars)%>%
  mutate(left_company = if_else(left_company == "Yes", 1, 0) )



# Rename Dummy variables as needed
employeeDF <- employeeDF %>% rename(
    "department_IT_Analytics" ="department_IT and Analytics",
    "department_Product_Development" = "department_Product Development",
    "job_level_Senior_Manager" = "job_level_Senior Manager",
    "job_level_Vice_President" = "job_level_Vice President",
    "job_satisfaction_Very_High" = "job_satisfaction_Very High",
    "performance_rating_Meets_Expectations" ="performance_rating_Meets Expectations",
    "performance_rating_Minimally_Effective" ="performance_rating_Minimally Effective",
    "performance_rating_Not_Effective" = "performance_rating_Not Effective",
    "performance_rating_Exceeds_Expectations" = "performance_rating_Exceeds Expectations",
    "department_Finance_and_Operations" ="department_Finance and Operations" 
)

# Add or remove variables as needed
employeeDF = select(employeeDF, c(-"job_satisfaction_Low",-"performance_rating_Minimally_Effective",
                                  -"department_Product_Development",-"job_level_Vice_President",
                                  -"performance_rating_Not_Effective", -"business_travel_None", 
                                  -"marital_status_Single") )
  
 colnames(employeeDF)


```

```{r}
### Generate training and validation datasets
# For the prediction, we generate the training and validation set using the "rsample" package.
# Generate the training and validation dataset set. 

set.seed(739)
employee_split <- initial_split(employeeDF, prop = 0.60)
employee_training <- employee_split %>% training()
employee_test <- employee_split  %>%  testing()
```

```{r}
neuralFunction <- function(input,s ){


 nn <- neuralnet(left_company ~. , data = employee_nn, hidden =  input )

 plot(nn, rep = "best")

 #print(" Training Graphs")
 #plotFunction(nn,employee_training, paste(s," Training"))

 print(" Validation Graphs")
 plotFunction(nn,employee_test, paste(s," Validation"))
}
```

```{r}
plotFunction <- function(inputNN, data, s ){
# Make predictions
  predictions <- predict(inputNN, data, type = "response")
  
  # Convert probabilities to binary predictions
  binary_predictions <- ifelse(predictions > 0.5, 1, 0)
  
  # Create confusion matrix
  confusion_matrix <- table(binary_predictions, as.numeric(data$left_company) )
  
  # Evaluate the model
  roc_curve <- roc(as.numeric(data$left_company) , as.numeric(predictions))
  NNtitle = paste( "AUC Curve:" ,s )
  plot(roc_curve, main = NNtitle, col = "red", lwd = 2)
  auc_score <- auc(roc_curve)
   
  # Print confusion matrix and AUC ROC
  print("Confusion Matrix:")
  print(confusion_matrix)
  
  cat("\n AUC ROC Score:", auc_score)
  cat( "\n" ,s)
  cat( "\n\n\n")
}
  
```

```{r}
library(neuralnet)
library("pROC")
### Generate the neural network
# Next we use the neuralnet command to generate the neural network. In this network, we
# specify hidden = 2, which corresponds to & hidden neurons in each lager. The model include
#all the dummies we'd generated earlier.

employee_nn <- employee_training

#employee_nn = select(employee_nn, -1)

hidden_df <- data.frame(
  hidden_values = I(list(c(5), c(5, 5), c(5, 5, 5)))
   # Test program:
   # hidden_values = I(list(c(1), c(1, 1), c(1, 1, 1)))
)

arrayString <- c("NN: 1 hidden layer with 5 nodes",
"NN: 2 hidden layer with 5 nodes",
"NN: 3 hidden layer with 5 nodes")


```

```{r}
## Loop over different Neural Networks

## • Neuralnet with one hidden layer with five nodes
## • Neuralnet with two hidden layers with five nodes in each layer
## • Neuralnet with three hidden layers and five nodes in each layer

for (index in 1:1) {
 
  cat("\n")
  neuralFunction( hidden_df$hidden_values[[index]], arrayString[index] )
  cat("\n")

}
```
\newpage

## Results

###  • Neuralnet with one hidden layer with five nodes

 Confusion Matrix:
 
         binary_predictions    0  1
 
                           0  489 12
                  
                           1   7  80

AUC ROC Score: 0.98054

### • Neuralnet with two hidden layers with five nodes in each layer

Confusion Matrix:
         
         binary_predictions     0   1

                           0   487 13
         
                           1   9  79
         
AUC ROC Score: 0.92571
         
### • Neuralnet with three hidden layers and five nodes in each layer

Confusion Matrix:

        binary_predictions       0   1

                             0  486 10
  
                             1  10  82

 AUC ROC Score: 0.9522923

## Conclusion

Neuralnet with one hidden layer with five nodes performed the best out of the three models with an  AUC ROC Score of 0.98054.

Neuralnet with three hidden layers and five nodes in each layer performed the second best out of the three models with an  AUC ROC Score of 0.9522923.
 
While, Neuralnet with two hidden layers with five nodes in each layer performed the worst out of the three models with an  AUC ROC Score of 0.92571


